---
title: "CSCI-89B Final Project"
subtitle: "The Computational Prometheus: Tracing the Evolution of Sci-Fi through Structural Topic Modeling"
author: "Liam Yardley"

output:
  pdf_document:
    toc: yes
    toc_depth: 2
  latex_engine: xelatex
  html_document:
    df_print: paged
---

# Setup

```{r setup, include=TRUE}
required_packages = c(
  # Data manipulation and plotting
  "tidyverse",
  # Downloading books from Project Gutenberg
  "gutenbergr", 
  # Tidy text mining principles
  "tidytext", 
  # Lemmatization tools
  "textstem", 
  # Structural Topic Modeling
  "stm",
  # Text Mining infrastructure
  "tm",
  # Stemming algorithms
  "SnowballC",
  # Topic modeling evaluation
  "textmineR", 
  # Word cloud visualization
  "wordcloud",   
  # String processing
  "stringi",
  # Graph plotting
  "igraph",  
  "gridExtra"
)

# Install missing packages automatically
new_packages = required_packages[!(required_packages %in% installed.packages()[,"Package"])]
if(length(new_packages)) install.packages(new_packages)

# --- Core Data Science Stack ---
library(tidyverse) 
library(stringi) 

# --- NLP & Text Mining Stack ---
library(gutenbergr) 
library(tidytext) 
library(textstem) 
library(tm)
library(SnowballC) 

# --- Modeling & Visualization Stack ---
library(stm)
library(textmineR) 
library(wordcloud) 
library(igraph)
library(gridExtra)
knitr::opts_chunk$set(echo = TRUE)
```

## Useful Functions
```{r functions}
# --- Pretty Tables ---
library(knitr)
create_table = function(data_list, 
                                row_labels = NULL, 
                                col_labels = NULL, 
                                transpose = FALSE, 
                                digits = 3,
                                caption = "Summary Table") {
  
  # Combine inputs into a data frame
  df = as.data.frame(data_list, stringsAsFactors = FALSE)
  
  # Assign row and column names
  if (!is.null(row_labels)) rownames(df) = row_labels
  if (!is.null(col_labels)) colnames(df) = col_labels
  
  # Optional transpose
  if (transpose) {
    df = as.data.frame(t(df))
  }

  # Apply numeric formatting
  df[] = lapply(df, function(x) {
      if (is.numeric(x)) {
        return(formatC(x, digits = digits, format = "fg", flag = "#"))
      } else {
        return(x)
      }
    })

  # Return nicely formatted knitr::kable
  knitr::kable(df, 
           
               booktabs = TRUE, 
               longtable = TRUE, 
               caption = caption, 
               position = "h!",
               row.names = FALSE)
}

# --- Printing ---
# These were removed after images were generated

# For ggplots
# ggsave("filename.pdf", 
#        plot = plot, 
#        width = 36, # Adjust as needed 
#        height = 12, # Adjust as needed 
#        units = "cm")
# 
# # For base R
# pdf("filename.pdf", width = 12, height = 8) # Adjust as needed (in inches)
# 
# # Plotting code here
# 
# dev.off()
```


#  Introduction

This workflow demonstrates the application of Structural Topic Modeling (STM) for literary genre classification. Using Mary Shelley's Frankenstein and The Last Man as test cases, we explore how probabilistic topic models can quantify the thematic distance between ambiguous texts and established genre categories.

Shelley's work occupies a unique position in literary history, often debated as either Romantic, Gothic Horror, or the progenitor of Science Fiction. To resolve this classification quantitatively, we construct a comparative corpus comprising:

- Realism/Romance: Representing the stylistic baseline of the Romantic/Victorian era.

- Gothic Horror: Representing themes of fear, the supernatural, and atmospheric dread.

- Science Fiction: Representing themes of technology, exploration, and future speculation.

# Objectives

This analysis proceeds in three stages:

Stylometric Comparison (PCA): Mapping the syntactic similarity between authors.

Topic Modeling (STM): Identifying latent thematic clusters (e.g., "Naval Exploration" vs. "Space Travel").

Genre Classification: Using topic prevalence to determine if Shelley's narrative structure aligns more closely with the static horror of the Gothic or the dynamic exploration of Science Fiction.

Note: While modern discussions often link Frankenstein to Artificial Intelligence (see Botting, 2017), this analysis focuses strictly on the historical genre evolution from 1800 to 1950.

# Data Handling
## Corpus Selection

The dataset consists of 44 novels selected from Project Gutenberg. To ensure a robust classification model, we define three "Control Groups" (Romance, Gothic, Sci-Fi) to act as the training context for our "Target Works" (Shelley).

The selection includes canonical texts chosen for their genre-defining characteristics. For instance, War of the Worlds and The Plague Ship are included specifically to provide a thematic baseline for the apocalyptic narratives found in Shelley's The Last Man.

The code below defines the Project Gutenberg IDs for the corpus and downloads the raw text data. These could be changed to adapt this code or genre classification of any works.

This data is available at Project Gutenberg (2025), https://www.gutenberg.org/.


```{r data-collection}
# --- Download Data ---
shelley_ids = c(
  # Target Works
  84, # Frankenstein by Mary Shelley
  18247 # The Last Man by Mary Shelley
)

romance_ids = c(
  # Romance
  1342, # Pride and Prejudice by Jane Austen
  161, #	Sense and Sensibility by	Jane Austen
  1260, # Jane Eyre by Charlotte Bronte
  768, # Wuthering Heights by Emily Brontë
  1400, # Great Expectations by Charles Dickens
  158, # Emma by Jane Austen
  730, # Oliver Twist by Charles Dickens
  98, # A Tale of Two Cities by Charles Dickens
  110, # Tesse of the d'Urbervilles by Thomas Hardy
  107, # Far from the Madding Crowd by Thomas Hardy
  514, # Little Women by Louisa May Alcott
  1399 # Anna Karenina by Leo Tolstoy
)

gothic_ids = c(
  # Gothic Horror
  345, # Dracula by Bram Stoker
  1188, #	The Lair of the White Worm by Bram Stoker
  3781, #	The Jewel of Seven Stars by Bram Stoker
  43, # Dr. Jekyll and Mr. Hyde by Robert Louis Stevenson
  174, # The Picture of Dorian Gray by Oscar Wilde
  10007, # Carmilla by J. Sheridan Le Fanu
  175, # Phantom of the Opera by Gaston Leroux
  2610, # The Hunchback of Notre-Dame by Victor Hugo
  6087, # The Vampyre by John Polidori
  696, #	The Castyle of Otranto	by Horace Walpole
  3268, # The Mysteries of Udolpho by Ann Radcliffe
  41,	# The Legend of Sleepy Hollow byWashington Irving
  583 # The Woman in White by Wilkie Collins
)

scifi_ids = c(
  35, # The Time Machine by HG Wells
  36, # The War of the Worlds by HG Wells
  159, # The Island of Doctor Moreau by HG Wells
  1013, # The First Men in the Moon by H.G. Wells
  5230, # The Invisible Man by H.G. Wells
  164, # 20,000 Leagues Under the Sea by Jules Verne
  18857, # Journey to the Center of the Earth by Jules Verne
  139, # The Lost World by Arthur Conan Doyle
  126, # The Poison Belt by Arthur Conan Doyle
  32032, # Second Variety by Philip K.Dick
  32154, # The Variable Man by Philip K.Dick
  16921, # Plague Ship by Andre Norton
  20782, # Triplanetary by E.E. "Doc" Smith
  32, # Herland by Charlotte Perkins Gilman
  21970, # The Scarlet Plague by Jack London
  32530 # Armageddon 2419 A.D. by Philip Francis Nowlan
)

book_ids = c(shelley_ids, romance_ids, gothic_ids, scifi_ids)

# --- Get Books ---
mirror = "http://mirrors.xmission.com/gutenberg/"

raw_books = gutenberg_download(book_ids, meta_fields = c("title", "author"), mirror = mirror)

# --- Clean and Label ---
books_clean = raw_books %>%
  # Remove Gutenberg Licence info
  filter(!is.na(text), text != "") %>%
  mutate(
    # Assign GENRE
    genre = case_when(
      gutenberg_id %in% shelley_ids ~ "Shelley",
      gutenberg_id %in% romance_ids ~ "Realism/Romance",
      gutenberg_id %in% gothic_ids  ~ "Gothic Horror",
      gutenberg_id %in% scifi_ids   ~ "Sci-Fi",
      TRUE ~ "Other"
    ),
    
    # Assign ERA (Approximate historical groups)
    era = case_when(
      # Modern Sci-Fi (1930s-1950s)
      gutenberg_id %in% c(32032, 32154, 16921, 20782) ~ "Modern (Atomic Age)",
      
      # Late Victorian(1880s-1920s)
      gutenberg_id %in% c(35, 36, 159, 139, 8748, 32, 345, 1188, 3781, 174, 43, 175, 110, 107) ~ "Late Victorian (1880-1920)",
      
      # Default to Romantic / Early-Mid Victorian (1800s-1870s)
      TRUE ~ "Romantic/Mid-Victorian (Pre-1880)"
    ),
  )

# --- Book Summary ---
book_summary = books_clean %>%
  select(genre, title, author, era, gutenberg_id) %>%
  distinct(gutenberg_id, .keep_all = TRUE) %>%
  arrange(genre, author)

create_table(book_summary, 
             caption = "Corpus of Selected Texts", 
             col_labels=c("Genre", "Book Title", "Author", "Era", "ID"))
```

Structural Topic Models require a high volume of discrete data points to detect patterns effectively. Since a novel constitutes a single, lengthy document, we must segment the text to increase our sample size. These works will be chunked into 100 words, to allow for sufficient data for Structural Topic Modelling. This has created a total of $42,204$ documents as my corpus.
Below is a sample of what these chunks would look like.


```{r data-processing}
# --- Chunking ---
documents_df = books_clean %>%
  unnest_tokens(word, text) %>%
  # Lemmatise (instead of Stemming)
  mutate(word = lemmatize_words(word)) %>%
  group_by(title) %>%
  # Create chunks of 100 words
  mutate(chunk_id = row_number() %/% 100) %>%
  # Ungroup first to ensure metadata carries over correctly
  ungroup() %>%
  group_by(title, genre, era, chunk_id) %>%
  summarize(text_segment = paste(word, collapse = " "), .groups = "drop") %>%
  # Create Unique Doc IDs
  mutate(doc_id = paste(str_sub(title, 1, 10), chunk_id, sep = "_"))

# --- Check Results ---
print(paste("Total Documents Created:", nrow(documents_df)))

create_table(table(documents_df$genre), caption = "Number of Documents per Genre", col_labels=c("Genre", "Document Count"))

create_table(head(documents_df,10), caption = "Sample of Documents")
```

A preliminary check of the document counts reveals a significant class imbalance. The Romantic novels in this corpus are substantially longer than the Gothic or Sci-Fi works, resulting in a disproportionate number of text chunks. To prevent the model from overfitting to the dominant class (Realism/Romance), we implement a random down-sampling strategy to cap each genre at 8,000 documents.

Note: This step is specific to the distribution of this particular corpus. If replicating this analysis with a smaller dataset, this block should be modified or omitted to avoid discarding data.

```{r data-sampling}
# --- Balance Corpus by Sampling Genres ---
set.seed(42)
target = 8000

balanced_df = documents_df %>%
  group_by(genre) %>%
  slice_sample(n = target) %>%
  ungroup()

df_original = as.data.frame(table(documents_df$genre))
colnames(df_original) = c("Genre", "Original_Count")

df_balanced = as.data.frame(table(balanced_df$genre))
colnames(df_balanced) = c("Genre", "Sample_Count")

sampling_summary = merge(df_original, df_balanced, by = "Genre")

# Add a column showing the Percentage Kept
sampling_summary$Percent_Kept = paste0(
  round((sampling_summary$Sample_Count / sampling_summary$Original_Count) * 100, 1), 
  "%"
)

create_table(sampling_summary, caption = "Number of Documents per Genre After Sampling", col_labels=c("Genre", "Original Count", "Sample Count", "Percent Kept"))
```


## Data Preparation

Before conducting STM, we need to preprocess the data I will be using the in-built `textProcessor`, and I will be removing stopwords, numbers, punctuation and stemming the words, as well as restricting to stems of a length $3$ or more.

Initially used stemming, this did not yield nice, interpretable words, which is especially important when analysing literature and considering key themes and topics.

### Consideration of Stop Words

One issue that may arise (and did on my first attempt) is very common words dominating topics, but are ultimately meaningless. For example, *chapter*, yes, this could have some meaning, like the 'chapter of my life', but that would be a rare occurrence, instead this word will exist dozens or more times in each novel, but serve no purpose. Therefore, I will create a custom set of stop words to eliminate these.

Another issue I came across, was the STM model being lazy and creating topics that were just books. As we have chunked each novel into 100 word chunks, there is a lot of data for each book, and therefore the model can just group these chunks into topics based on the book they came from, rather than any thematic or topical similarity. To counteract this, I created a custom stop word list that included book titles, author names, and other common words that were not useful for topic modelling. One major think that could impact that would be names, therefore I have added the main character names from each book to the stop word list to try and prevent this issue (names generated with help from AI: Gemini)

```{r custom-stopwords}
# --- Custom Stopwords ---
# Generation of names to be omitted supported by Gemini
my_stopwords = c(
  # --- Shelley ---
  # Frankenstein
  "victor", "frankenstein", "elizabeth", "lavenza", "clerval", "henry", "walton", 
  "robert", "justine", "moritz", "felix", "agatha", "safie", "delacey", "creature", "monster",
  # The Last Man
  "lionel", "verney", "adrian", "raymond", "perdita", "idris", "evadne", "windsor",
  
  # --- Realism & Romance ---
  # Austen (P&P, S&S, Emma)
  "darcy", "bennet", "bingley", "wickham", "collins", "lydia", "kitty", "gardiner",
  "elinor", "marianne", "dashwood", "edward", "ferrars", "willoughby", "brandon",
  "emma", "woodhouse", "knightley", "churchill", "fairfax", "bates", "elton",
  # Bronte Sisters (Jane Eyre, Wuthering Heights)
  "jane", "eyre", "rochester", "adele", "varens", "bertha", "mason", "rivers", "temple",
  "heathcliff", "catherine", "cathy", "earnshaw", "linton", "lockwood", "nelly", "dean",
  # Dickens (Great Expectations, Oliver Twist, Tale of Two Cities)
  "pip", "pirrip", "estella", "havisham", "gargery", "magwitch", "jaggers", "drummle",
  "oliver", "twist", "fagin", "sikes", "nancy", "bumble", "brownlow", "dodger",
  "carton", "darnay", "manette", "lorry", "defarge", "pross", "stryver",
  # Hardy (Tess, Far From Madding Crowd)
  "tess", "durbeyfield", "angel", "clare", "alec", "durberville", 
  "bathsheba", "everdene", "gabriel", "oak", "troy", "boldwood",
  # Other Classics (Little Women, Anna Karenina)
  "jo", "meg", "beth", "amy", "march", "laurie", "laurence", "bhaer",
  "anna", "karenina", "vronsky", "levin", "kitty", "stepan", "stiva", "oblonsky", "dolly", "alexey",
  
  # --- Gothic Horror ---
  # Stoker (Dracula, Lair, Jewel)
  "dracula", "helsing", "harker", "jonathan", "mina", "seward", "holmwood", "godalming", "renfield",
  "adam", "salton", "arabella", "caswall", "oolanga", "mim", 
  "malcolm", "ross", "trelawny", "tera", "corbeck", "winchester",
  # Wilde, Stevenson, Le Fanu
  "dorian", "gray", "basil", "hallward", "wotton", "sybil", "vane",
  "jekyll", "hyde", "utterson", "lanyon", "enfield", "poole",
  "carmilla", "laura", "spielsdorf", "karnstein",
  # Leroux, Hugo, Polidori
  "phantom", "erik", "christine", "daae", "raoul", "chagny", "giry",
  "quasimodo", "esmeralda", "frollo", "phoebus", "gringoire",
  "ruthven", "aubrey", "ianthe",
  # Walpole, Radcliffe, Irving, Collins
  "manfred", "conrad", "isabella", "theodore", "matilda", "jerome", "otranto",
  "emily", "aubert", "valancourt", "montoni", "udolpho", "laurentini",
  "ichabod", "crane", "katrina", "tassel", "brom", "bones",
  "walter", "hartright", "marian", "halcombe", "fosco", "percival", "glyde", "catherick",
  
  # --- Sci-Fi ---
  # Wells (Time Machine, War of Worlds, Moreau)
  "weena", "morlock", "eloi", "traveller", "ogilvy", "henderson",
  "prendick", "moreau", "montgomery",
  # Verne (20k Leagues, Journey)
  "nemo", "nautilus", "aronnax", "conseil", "ned", "land", 
  "lidenbrock", "axel", "hans", "grauben",
  # Doyle, Burroughs (Lost World, Mars)
  "challenger", "malone", "roxton", "summerlee",
  "john", "carter", "dejah", "thoris", "tars", "tarkas", "sola", "woola", "barsoom",
  # Dick, Norton, Smith, Gilman (Modern Era)
  "hendricks", "tasso", "klaus", "rudi", "david", 
  "cole", "reinhart", "kaplan", "sherikov",
  "dane", "thorson", "rip", "ali", "kamil", "jellico", "tau", "sargol",
  "costigan", "clio", "bradley", "rog", "nerado", "samms", "kinnison",
  "van", "vandyck", "terry", "jeff", "ellador", "celia", "alima", "herland",
  
  # --- Procedural Junk ---
  "chapter", "vol", "volume", "book", "part", "gutenberg", "project", "edition", 
  "contents", "title", "author", "prologue", "epilogue", "illustration"
)
```

```{r stm-modeling}
# --- Prepare Data for STM ---
processed = textProcessor( 
  documents = balanced_df$text_segment,   
  metadata = balanced_df[, c("title", "genre", "era")],
  lowercase = TRUE,
  removestopwords = TRUE,
  removenumbers = TRUE,
  removepunctuation = TRUE,
  stem = FALSE,
  customstopwords = my_stopwords,
  wordLengths = c(3, Inf),
  verbose = TRUE
)

# Build documents, vocabulary, and aligned metadata
out = prepDocuments(processed$documents, processed$vocab, processed$meta, lower.thresh = 10)
docs = out$documents
vocab = out$vocab
meta = out$meta

# --- Summaries ---
num_docs = length(docs)
vocab_size = length(vocab)
meta_dim = dim(meta)

cat("Number of documents:", num_docs, "\n\n")
cat("Vocabulary size:", vocab_size, "\n\n")
cat("Metadata dims (rows, cols):", meta_dim[1], meta_dim[2], "\n\n")

# Average document length (in tokens) after preprocessing:
avg_len = mean(sapply(docs, length))
cat("Avg tokens per document (post-preprocessing):", round(avg_len, 2), "\n\n")

# Check
head(vocab, 20)
head(meta)

# Example preprocessed document
cat("\nExample preprocessed document (first 20 tokens):\n")
print(docs[[1]][1:20])
```

# Style Comparison

Beyond thematic content, we can analyze the *stylistic* fingerprints of the authors. Style is often defined not by unique nouns (content), but by the distribution of high-frequency function words (the, and, of, but).

The following code block prepares the data for Principal Component Analysis (PCA) by:
1.  Isolating the top 150 most frequent words in the entire corpus.
2.  Calculating the relative frequency of these words for each book.
3.  Projecting this high-dimensional data onto a 2D plane to visualize stylistic clusters.

```{r pca}
TOP_N = 150

# --- Prepare Style ---
# We go back to the raw books
style_dtm = books_clean %>%
  unnest_tokens(word, text) %>%
  
  # Count word usage per book
  count(title, word) %>%
  group_by(title) %>%
  mutate(total_words = sum(n)) %>%
  
  # Calculate Relative Frequency (Frequency per 1000 words)
  # Normalise for book length
  mutate(freq = (n / total_words) * 1000) %>%
  ungroup() %>%
  
  # Filter for the Top N Most Frequent Words
  group_by(word) %>%
  mutate(word_total = sum(n)) %>%
  ungroup() %>%
  filter(word %in% slice_max(unique(select(., word, word_total)), order_by = word_total, n = TOP_N)$word) %>%
  
  # Reshape to Matrix (Rows = Books, Cols = Words)
  select(title, word, freq) %>%
  pivot_wider(names_from = word, values_from = freq, values_fill = 0)

# --- PCA ---
# Convert to matrix
pca_matrix = as.matrix(style_dtm[, -1])
rownames(pca_matrix) = style_dtm$title

# Run PCA
pca_result = prcomp(pca_matrix, scale. = TRUE)

# --- Plot the Literary Map ---
# Extract coordinates
pca_coords = as.data.frame(pca_result$x)
pca_coords$title = rownames(pca_coords)

# Add Genre back in for coloring
pca_plot_data = pca_coords %>%
  left_join(unique(books_clean[, c("title", "genre")]), by = c("title" = "title"))

# Visualize
my_plot = ggplot(pca_plot_data, aes(x = PC1, y = PC2, color = genre, label = title)) +
  geom_point(size = 4, alpha = 0.8) +
  geom_text(vjust = 1.5, size = 3, check_overlap = FALSE) +
  labs(title = "Stylometric Map Corpus",
       subtitle = "Based on usage of top 150 function words (PCA)",
       x = "PC1 (Primary Stylistic Dimension)",
       y = "PC2 (Secondary Stylistic Dimension)") +
  theme_minimal()

ggsave("Stylometric Map.pdf", 
       plot = my_plot, 
       width = 36, 
       height = 24, 
       units = "cm")
```

The PCA plot reveals that Gothic Horror is a stylistically diverse genre. Some works cluster near Realism/Romance, while others approach Science Fiction. Notably, early Gothic works like The Castle of Otranto and The Mysteries of Udolpho appear distinct from later Victorian Gothic novels like Dracula.

Mary Shelley's works lie in a unique neighborhood, stylistically adjacent to early Science Fiction works by Jules Verne. This suggests a potential stylistic lineage, though definitive influence is difficult to prove from clustering alone.

Cosine Similarity Analysis
PCA provides a useful 2D approximation, but it inherently loses information by compressing the dimensions. To get a more precise measure of similarity, we calculate the Cosine Similarity between the full PCA vectors of Shelley's works and every other book in the corpus.


```{r closeststyle}
# --- Find Closest Books in Style ---
# Cosine Similarity Function
cosine_sim = function(a, b) {
  sum(a * b) / (sqrt(sum(a^2)) * sqrt(sum(b^2)))
}

# --- Frankenstein ---

target_vec = pca_matrix["Frankenstein; Or, The Modern Prometheus", ]

similarities = data.frame(
  title = rownames(pca_matrix),
  similarity = apply(pca_matrix, 1, function(x) cosine_sim(target_vec, x))
)

# Get Top 10 Most Similar Books
top_similar = similarities %>%
  filter(title != "Frankenstein; Or, The Modern Prometheus") %>%
  arrange(desc(similarity)) %>%
  slice_head(n = 10) %>%
  left_join(unique(books_clean[, c("title", "genre")]), by = c("title" = "title"))

create_table(top_similar, 
             caption = "Top 10 Books with a similar style to Mary Shelley's Frankenstein", 
             col_labels=c("Book Title", "Cosine Similarity"))

top_similar %>%
  ggplot(aes(x = reorder(title, similarity), y = similarity, fill = genre)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Books with a similar style to Mary Shelley's 'Frankenstein'",
    subtitle = "Top 10 Stylistic Matches (Cosine Similarity)",
    y = "Cosine Similarity Score",
    x = NULL,
    fill = "Genre"
  ) +
  theme_minimal() +
  scale_y_continuous(limits = c(0.95, 1.0), oob = scales::rescale_none)



# --- The Last Man ---

target_vec = pca_matrix["The Last Man", ]

similarities = data.frame(
  title = rownames(pca_matrix),
  similarity = apply(pca_matrix, 1, function(x) cosine_sim(target_vec, x))
)

# Get Top 10 Most Similar Books
top_similar = similarities %>%
  filter(title != "The Last Man") %>%
  arrange(desc(similarity)) %>%
  slice_head(n = 10) %>%
  left_join(unique(books_clean[, c("title", "genre")]), by = c("title" = "title"))

create_table(top_similar, 
             caption = "Top 10 Books with a similar style to Mary Shelley's 'The Last Man'", 
             col_labels=c("Book Title", "Cosine Similarity"))

top_similar %>%
  ggplot(aes(x = reorder(title, similarity), y = similarity, fill = genre)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Books with a similar style to Mary Shelley's 'The Last Man'",
    subtitle = "Top 10 Stylistic Matches (Cosine Similarity)",
    y = "Cosine Similarity Score",
    x = NULL,
    fill = "Genre"
  ) +
  theme_minimal() +
  scale_y_continuous(limits = c(0.95, 1.0), oob = scales::rescale_none)
```

This analysis indicates that Frankenstein sits on the stylistic border between Gothic Horror and Science Fiction. If we were to use a K-Nearest Neighbors approach ($k=5$), the classification would oscillate between the two genres depending on the specific value of $k$.Interestingly, The Last Man shares strong stylistic similarities with much later Science Fiction works, despite its early publication date. This supports the hypothesis that Shelley was not merely writing within the Romantic tradition, but was actively pioneering a new stylistic register that would later define speculative fiction.Having established the stylistic context, the next section applies Structural Topic Modeling to analyze the thematic content of the works.


# Structural Topic Modelling

## Number of Topics

As the corpus of text is very large, we would expect for there to be a large number of topics. Therefore, using `searchK` will aim to find the optimal number of topics for my analysis. First, we will do a wide search of $K\in \{15,20,25,30,35,40,45,50\}$.

```{r searchk}
# --- Finding optimal number of topics --
# Set seed for reproducibility
set.seed(42)

# Use searchK to evaluate models with different numbers of topics
k_result = searchK(
  documents = docs,
  vocab = vocab,
  K = c(15,20,25,30,35,40,45,50),
  prevalence = ~ genre + era,
  data = meta,
  max.em.its = 150,
  init.type = "Spectral",
  verbose = FALSE)

plot(k_result)
```

```{r searchkplot1}
# ---Plotting Coherence vs. Exclusivity ---
k_values = unlist(k_result$results$K)

coherence = sapply(k_result$results$semcoh, function(x) x[1])
exclusivity = sapply(k_result$results$exclus, function(x) x[1])

plot_data = data.frame(Topics = k_values, Coherence = coherence, Exclusivity = exclusivity)

ggplot(plot_data, aes(x = Exclusivity, y = Coherence, label = Topics)) +
  geom_point() +
  geom_text(vjust = -0.5) +
  labs(title = "Exclusivity vs. Coherence", x = "Exclusivity", y = "Semantic Coherence") +
  theme_minimal()
  
```

In order to get a balance of coherence and exclusivity, I will scale the these values, and construct a composite score for easier identification of the optimal number of topics. I will do this using the following

$$\text{Composite Score} = 0.5 \times \text{Coherence} + 0.5\times \text{Exclusivity}$$

```{r searchkplot2}
# --- Graphing Composite Score ---
rescale = function(x) {
  (x - min(x)) / (max(x) - min(x))
} 

rescaled_coh = rescale(coherence)
rescaled_exc = rescale(exclusivity)
rescaled_data = data.frame(Topics = k_values, Rescaled_Coherence = rescaled_coh, Rescaled_Exclusivity = rescaled_exc)

create_table(rescaled_data, caption = "Rescaled Coherence and Exclusivity Scores")

# Calculate Composite Score
composite_score = 0.5 * rescaled_coh + 0.5 * rescaled_exc
composite_data = data.frame(Topics = k_values, Composite_Score = composite_score)

create_table(composite_data, caption = "Composite Scores for Different Numbers of Topics")

optimal_k = composite_data$Topics[which.max(composite_data$Composite_Score)]

cat("\nMaximum Composite Score at K =", optimal_k, "\n")

# Plot Composite Score vs. Number of Topics
ggplot(composite_data, aes(x = Topics, y = Composite_Score)) +
  geom_line() +
  geom_point() +
  labs(title = "Composite Score vs. Number of Topics", x = "Number of Topics", y = "Composite Score") +
  theme_minimal()
```


This gives an optimal $K \approx 35$. I initially wanted to hone in on the values around $35$ to find a better value of $K$, however, I am not sure of the impact this would have, especially with the time taken to run these searches. Also, in this exploration, a mathematically optimal value of $K$ may not be what is best for classifying these works, and we may need more topics to see the nuances between key themes.


## STM Model

From the previous section, I will fit an STM model with $K=35$ topics, and using `genre` and `era` as prevalence covariates. We will now run our model to find the $35$ topics for our corpus.

The top words generated for each topic will assist in classifying the topics into specific genres or themes for analysis.

```{r stm-fitting}
# --- Fit STM ---
set.seed(42)
stm_model = stm(
  documents = docs,
  vocab = vocab,
  K = optimal_k, # Number of topics
  prevalence = ~ genre + era,
  max.em.its = 150,
  data = meta,
  init.type = "Spectral",
  verbose = FALSE
)

# Number of Successful Runs
length(stm_model$runout)
```


```{r stm-analysis1}
# Top 7 words for each topic
labelTopics(stm_model, n=7)
```

## Topic Summary

Considering both the most common words,and the most exlusive can assist in our interpertion of each topic. If some words only appear in a topic, they much be significant to its interpretation.

```{r topic-analysis2}
label_output = labelTopics(stm_model, n = 5)

topic_labels_df = data.frame(
  Topic_ID = 1:nrow(label_output$prob),
  Top_Words_Prob = apply(label_output$prob, 1, paste, collapse = ", "),
  Unique_Words_FREX = apply(label_output$frex, 1, paste, collapse = ", ")
)

knitr::kable(topic_labels_df, 
      # format = "latex", 
      booktabs = TRUE, 
      longtable = TRUE,
      caption = "Topic Definitions: Probability vs. Exclusivity",
      col.names = c("Topic ID", "Highest Probability Words", "Most Exclusive (FREX) Words"),
      row.names = FALSE)
```

This considers the most common topics across the whole corpus. 

```{r topicleaderboard}
# Top words for each topic
td_beta = tidy(stm_model, matrix = "beta")

# For each document, proportion of each topic
td_gamma = tidy(stm_model, matrix = "gamma")

# Get top terms for each topic
top_terms = td_beta %>%
  group_by(topic) %>%
  slice_max(beta, n = 10) %>%  # Get the top 10 words per topic
  select(topic, term) %>%
  summarise(terms = list(term)) %>% # Collapse them into a list
  mutate(terms = map_chr(terms, paste, collapse = ", ")) # Paste them: "ship, sea, captain"

# Calculate average topic prevalence across all documents
gamma_terms = td_gamma %>%
  group_by(topic) %>%
  summarise(gamma = mean(gamma)) %>% # Calculate average prevalence
  arrange(desc(gamma)) %>%
  left_join(top_terms, by = "topic") %>% # Attach the word labels
  mutate(topic = paste0("Topic ", topic),
         topic = reorder(topic, gamma)) # Sort bars by size

# Plot the Top Topics by Prevalence
top_topics = gamma_terms %>%
  slice_max(gamma, n = 35) %>% 
  ggplot(aes(topic, gamma, label = terms, fill = topic)) +
  geom_col(show.legend = FALSE) +
  geom_text(hjust = 0, nudge_y = 0.0005, size = 3) + # Print the words next to bars
  coord_flip() + # Make it horizontal
  scale_y_continuous(expand = c(0,0), limits = c(0, 0.1)) + # Fix axis size
  labs(x = NULL, y = "Topic Prevalence",
       title = "Topics by Prevalence",
       subtitle = "With top words contributing to each topic")

ggsave("Top Topics.pdf", 
       plot = top_topics, 
       width = 18, # Adjust as needed 
       height = 12, # Adjust as needed 
       units = "cm")
```

## Topic Examples

In order to allow us to get a greater insight into each topic, we will consider the top $5$ examples of chunks of text that relate to these topics. This should help us determine what themes these topics are portraying, and to check our interpretation of them.

```{r topicexamples}
# --- Qualitative Audit ---
for (t in c(16,18)) {
  thoughts = findThoughts(
    stm_model, 
    texts = balanced_df$text_segment,
    topics = t,         
    n = 5
  )$docs[[1]]

  title = paste("Topic", t, "Examples")
  plotQuote(thoughts, width = 80, main = title)
}
```

# Topic Visualisation and Analysis

Now, the issue we come to now, is that we need to make some sense of these 'Topics' created by my STM model. The most frequent words, and the FREX (most unique) words will give some guide towards what these topics are, and the examples from each topic will also allow us to see the context of these topics. However, with $35$ topics, and with the possible nuances of each of them, this alone is not enough to fully understand their meaning. Therefore I need a wide range of different tools and visualisations to be able to interpret them for comparison. Therefore my final analysis of these topics will be at the end of this section, having looked through a range of visualisation. I will be considering

- **Top Terms per Topic**: To help visualise the most frequent words for each topic
- **Word Clouds**: Another way to visualise the most frequent words for each topic
- **Control Checks**: To see which topics align with specific texts to help compare topics, and also check that our topics, and our generalisation of them, do actually make sense.
- **Topic Prevalence by Genre**: To see which topics are more prevalent in which genres, and see if this aligns with our interpretations of the topics.
- **Topic Correlations**: To see which topics are related to each other, and see if this aligns with our interpretations of the topics, and to help group similar topics together.

Only all of this together will give us a full understanding and help to truly classify these topics and Shelley's works.

Once we have fully dissected our topics, we can then do **Topic Comparison** to compare specific topics between Shelley and different genres to see which she most aligns with.

## Top Terms per Topic

```{r stm-visualization1, fig.width=10, fig.height=18}
# Top terms per topic for visualization
# (From Julia Stilge's example)
beta_td = tidy(stm_model)

top_terms = beta_td %>%
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)


top_terms %>%
  mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free_y") +
  scale_y_reordered() +
  labs(title = "Top Terms per Topic",
       x = "Beta (word probability)",
       y = NULL)


```
## Word Clouds

```{r topicwordcloud, fig.width=10, fig.height=18, warning=FALSE}
library(RColorBrewer)

# choose a base palette with 3 distinct hues (for 3 topics)
base_cols = brewer.pal(3, "Dark2") 

# Word clouds for each topic
par(mfrow = c(9, 4), mar = c(0,0,1,0))
for (k in 1:optimal_k) {
  cloud(stm_model,
        topic = k,
        max.words = 25,
        scale = c(1.8, 0.4),
        colors = brewer.pal(8, "Dark2"))
  
  # add title after the cloud plot
  mtext(paste("Topic", k), line = -0.5, cex = 1.2, font = 2)
}
par(mfrow = c(1, 1))

```
## Control Checks

To ensure that our topics actually make sense in our context, as well as gain a deeper insight into each topic, we will check back to our novels, and make sure see the topics that are prevalent in some example of books. We expect to see to *Romance* themed topics (such as domestic life, romance, family) in *Pride and Prejudice*, *Gothic Horror* topics (such as supernatural, fear, death) in *Dracula*, and *Science Fiction* topics (such as technology, exploration, future) in *The Time Machine*. We will check this with a bar chart of prevalence of each topic, and see if it matches our previous interpretation of these topics.

```{r controlcheck}
# --- Topic Prevalence Check ---
topic_props = as.data.frame(stm_model$theta)
colnames(topic_props) = paste0("Topic_", 1:ncol(topic_props))

# Add metadata
meta_aligned = balanced_df 
if(length(out$docs.removed) > 0) {
  meta_aligned = meta_aligned[-out$docs.removed, ]
}

# Combine them
doc_topic_df = cbind(meta_aligned, topic_props)

# Define colors for genres
genre_colors = c(
  "Realism/Romance" = "#D32F2F",  # Deep Red
  "Gothic Horror"   = "#7B1FA2",  # Deep Purple
  "Sci-Fi"          = "#1976D2",  # Strong Blue
  "Shelley"         = "#FBC02D"  # Gold/Dark Yellow (Readable)
)

# Control Check Function
plot_book_topics = function(book_title_pattern, color_tone = "steelblue") {
  
  # Filter for the specific book
  book_data = doc_topic_df %>%
    filter(str_detect(title, book_title_pattern)) %>%
    select(starts_with("Topic_")) %>%
    colMeans() %>% 
    as.data.frame() %>%
    rename(Prevalence = ".") %>%
    mutate(Topic = rownames(.)) %>%
    arrange(desc(Prevalence)) %>%
    slice_head(n = 10)
  
  # Plot
  ggplot(book_data, aes(x = reorder(Topic, Prevalence), y = Prevalence, fill = Prevalence)) +
    geom_col(show.legend = FALSE) +
    coord_flip() +
    scale_fill_gradient(low = "grey90", high = color_tone) +
    labs(title = paste("Top Topics in:", book_title_pattern),
         y = "Average Probability", x = "Topic") +
    theme_minimal()
}

  
wrap_title = function(text) {
  str_wrap(text, width = 25) # Break line every 25 chars
}

# --- Romance ---
c_rom = genre_colors["Realism/Romance"]
p1 = plot_book_topics("Pride and Prejudice", c_rom) + labs(title=wrap_title("Pride and Prejudice (Austen)"))
p2 = plot_book_topics("Jane Eyre", c_rom) + labs(title=wrap_title("Jane Eyre (Bronte)"))
p3 = plot_book_topics("Great Expectations", c_rom) + labs(title=wrap_title("Great Expectations (Dickens)"))

romance_grid = grid.arrange(p1, p2, p3, ncol = 3, top = "Genre Check: Realism & Romance")

ggsave("Romance Check.pdf", 
       plot = romance_grid, 
       width = 36, 
       height = 12,
       units = "cm")

# --- Gothic Horror ---
c_goth = genre_colors["Gothic Horror"]
p1 = plot_book_topics("Dracula", c_goth) + labs(title=wrap_title("Dracula (Stoker)"))
p2 = plot_book_topics("Dr. Jekyll", c_goth) + labs(title=wrap_title("The Strange Case of Dr. Jekyll and Mr. Hyde (Stevenson)"))
p3 = plot_book_topics("Dorian Gray", c_goth) + labs(title=wrap_title("The Picture of Dorian Gray (Wilde)"))

gothic_grid = grid.arrange(p1, p2, p3, ncol = 3, top = "Genre Check: Gothic Horror")

ggsave("Gothic Check.pdf", 
       plot = gothic_grid, 
       width = 36, 
       height = 12,
       units = "cm")

# --- Science Fiction ---
c_sci = genre_colors["Sci-Fi"]
p1 = plot_book_topics("Time Machine", c_sci) + labs(title=wrap_title("The Time Machine (Wells)"))
p2 = plot_book_topics("War of the Worlds", c_sci) + labs(title=wrap_title("The War of the Worlds (Wells)"))
p3 = plot_book_topics("Lost World",c_sci) + labs(title=wrap_title("The Lost World (Doyle)"))

scifi_grid = grid.arrange(p1, p2, p3, ncol = 3, top = "Genre Check: Science Fiction")

ggsave("SciFi Check.pdf", 
       plot = scifi_grid, 
       width = 36, 
       height = 12,
       units = "cm")

# --- Shelley ---
c_shel = genre_colors["Shelley"]
p1 = plot_book_topics("Frankenstein", c_shel) + labs(title=wrap_title("Frankenstein; Or, The Modern Prometheus"))
p2 = plot_book_topics("Last Man", c_shel) + labs(title=wrap_title("The Last Man"))

shelley_grid = grid.arrange(p1, p2, ncol = 2, top = "Target Works: Mary Shelley")

ggsave("Shelley Check.pdf", 
       plot = shelley_grid, 
       width = 36, 
       height = 12, 
       units = "cm")
```

Based on all of these representations, and the most frequent and exclusive words in each topic, we can highlight what idea each topic is protraying, as well as assosiate them with a genre. However, to further aid this, it may be useful to see if these topics are related to each other, therefore we will next consider topic correlation.

## Topic Correlation

Themes that are quite common in a lot of these books, may be connected, to mulitple other topics. Therefore, we will consider the correlation between topics to see if any of them stand out as unique, or to see if we can cluster some topics together based on their correlations.

The Networks constructed here can be hard to read an interpret, so the following was done to improve this
- Change cutoff for correlation to $0.1$ to reduce number of edges
- Topics were colour coded to easily see clusters
- Shelley's topics have a yellow boarder, to see overlap with other genres
- Any topics in multiple genres were colour blended to show this crossover
- A table was created to show only correlations over $0.2$ to check for the strongest relationships.

For adapting this model, it is key to change these topics based on earlier analysis of the topics. These have been changed from the report as that was generated on an earlier iteration of the model. The remaining code has been updated to fit within this data, and from interpretations of the previous prevalence bar charts.

This could be automatically linked by outputting the above code as dataframe instead and defining topics. However, as we are only using selected texts, not the whole corpus, and considering qualitatively what each topic would relate to, we have stuck with manual updating.

```{r topiccorrelation}
# --- Topic Colour coding ---
# Romance (Red)
romance_topics = c(2,6,24) 
# Gothic (Purple)
gothic_topics = c(10,15,5)
# Sci-Fi (Blue)
scifi_topics = c(25,29,11,14)
# Shelley (Yellow)
shelley_topics = c(9,21,3)

# Base colours
cols = c(
  Romance = "#D32F2F",
  Gothic  = "#7B1FA2",
  SciFi   = "#1976D2",
  Shelley = "#FBC02D",
  Neutral = "lightsteelblue2"
)

# Create vectors of length 35 (one for each topic)
v_fill   = rep(cols["Neutral"], 35)
v_border = rep("grey60", 35)
v_width  = rep(1, 35)

# Loop through 1:35 to handle crossovers
for (i in 1:35) {
  my_genres = c()
  
  if (i %in% romance_topics) my_genres = c(my_genres, cols["Romance"])
  if (i %in% gothic_topics)  my_genres = c(my_genres, cols["Gothic"])
  if (i %in% scifi_topics)   my_genres = c(my_genres, cols["SciFi"])
  
  # Logic:
  if (length(my_genres) == 1) {
    # Single Genre, use that colour
    v_fill[i] = my_genres[1]
  } else if (length(my_genres) > 1) {
    # Crossover, Blend the colors
    # We pick the midpoint
    v_fill[i] = colorRampPalette(my_genres)(3)[2] 
  }
}

# Change border for Shelley Topics
for (i in shelley_topics) {
  v_border[i] = cols["Shelley"]
  v_width[i]  = 4
  }

# --- Topic Correlation Plot ---
topic_corr = topicCorr(stm_model, cutoff = 0.1)

plot(topic_corr,
     vertex.color = v_fill,
     vertex.frame.color = v_border, 
     vertex.frame.width = v_width,
     vertex.label.cex = 0.7,
     vertex.label.color = "black",
     main = "Topic Correlation Network")

legend("topleft", 
       legend = c("Romance", "Gothic", "Sci-Fi", "Shelley (Border Only)", "Neutral", "Crossover"), 
       fill = c(cols["Romance"], cols["Gothic"], cols["SciFi"], "white", cols["Neutral"], "white"), 
       border = c("black", "black", "black", cols["Shelley"], "black", "white"),
       pt.lwd = c(1,1,1,3,1),
       bty = "n", cex = 0.8)

# --- Strong correlations ---
cor_matrix = topic_corr$cor

# Remove self-correlations
diag(cor_matrix) = NA

# Convert to a dataframe (Topic A | Topic B | Correlation)
cor_df = as.data.frame(cor_matrix) %>%
  mutate(Topic_A = row_number()) %>%
  pivot_longer(
    cols = -Topic_A, 
    names_to = "Topic_B_Label", 
    values_to = "Correlation"
  ) %>%
  # Extract the number from "V11" -> 11
  mutate(Topic_B = as.integer(str_extract(Topic_B_Label, "\\d+"))) %>%
  filter(!is.na(Correlation)) %>% 
  filter(Topic_A < Topic_B) %>%
  arrange(desc(Correlation))

# Filter and Clean up columns
strong_links = cor_df %>%
  filter(Correlation > 0.2) %>%
  mutate(
    Topic_A = as.integer(Topic_A),
    Topic_B = as.integer(Topic_B),
    Correlation = round(Correlation, 3)
  ) %>%
  # --- THE FIX: Select only what you want, in the order you want ---
  select(Topic_A, Topic_B, Correlation)

# Print the clean table
knitr::kable(strong_links, caption="Strong Topic Correlations (r > 0.2)")
```

From here we can see some clear groupings and clusters between different genres.

## Topic Prevalence by Genre

Key graph to show the range of prevalence of each topic in the four genres. This will help us see which topics are more prevalent in which genres, and see if this aligns with our interpretations of the topics.This allows us to see where Shelley's works sit in comparison to the other genres, and where she works may have influenced or been influenced by other genres.

To be able to interpret this well, we have seperated topics by the genre they assosiate with, from out previous methods. Therefore, we can see which topics from each genre are more prevalent in Shelley's works, and which topics from other genres are less prevalent. This will help us see which genre Shelley's works align with the most. 

```{r topicprevalence}
# --- Topic Prevalence by Genre ---
topic_effect = estimateEffect(
  1:optimal_k ~ genre + era, 
  stm_model, 
  meta = out$meta)
```

```{r topicprevalence-plot}
# Romance (Red)
romance_topics = c(2,6,24) 
# Gothic (Purple)
gothic_topics = c(10,15,5)
# Sci-Fi (Blue)
scifi_topics = c(25,29,11,14)
# Shelley (Yellow)
shelley_topics = c(9,21,3)

# --- Romance Prevalence ---
par(mfrow = c(2,2))
for (topic in romance_topics) {
  
  plot(
    topic_effect,
    covariate = "genre",
    topics = topic,
    model = stm_model,
    method = "pointestimate",
    xlab = "Genre",
    
    main = paste("Topic", topic, "Prevalence"), 
    
    labeltype = "custom",
    custom.labels = c("Gothic", "Romantic", "Sci-Fi", "Shelley") 
  )
}


# --- Gothic Prevalence ---
par(mfrow = c(2,2))
for (topic in gothic_topics) {
  
  plot(
    topic_effect,
    covariate = "genre",
    topics = topic,
    model = stm_model,
    method = "pointestimate",
    xlab = "Genre",
    
    main = paste("Topic", topic, "Prevalence"), 
    
    labeltype = "custom",
    custom.labels = c("Gothic", "Romantic", "Sci-Fi", "Shelley") 
  )
}


# --- Science Fiction Prevalence ---
par(mfrow = c(2,2))
for (topic in scifi_topics) {
  
  plot(
    topic_effect,
    covariate = "genre",
    topics = topic,
    model = stm_model,
    method = "pointestimate",
    xlab = "Genre",
    
    main = paste("Topic", topic, "Prevalence"), 
    
    labeltype = "custom",
    custom.labels = c("Gothic", "Romantic", "Sci-Fi", "Shelley") 
  )
}

# --- Shelley Prevalence ---
par(mfrow = c(2,2))

for (topic in shelley_topics) {
  
  plot(
    topic_effect,
    covariate = "genre",
    topics = topic,
    model = stm_model,
    method = "pointestimate",
    xlab = "Genre",
    
    main = paste("Topic", topic, "Prevalence"), 
    
    labeltype = "custom",
    custom.labels = c("Gothic", "Romantic", "Sci-Fi", "Shelley") 
  )
}

# Reset layout
par(mfrow = c(1, 1))
```

## Topic Comparison

This visualisation allows us to see the comparison between two topics, and see which words are more associated with each topic. Therefore, we can compare Shelley's topics to those of other genres, and see which words are more associated with each genre. This will help us see which genre Shelley's works align with the most. 

These topics need to be picked based on previous analysis. The best way to utilise this is to consider two topics that are related between two genres, but are possibly dealt with differently. For example, we considered topics related to explaining scenery using different topics, or love in order to compare their different treatment in each genre.


```{r genrevisualisation, warning=FALSE}
# --- Genre Visualisation ---
# "Shelley" vs. "Realism/Romance"
plot(
  stm_model, 
  type = "perspectives", 
  topics = c(9,6), 
  model = stm_model, 
  cov.value1 = "Shelley",
  cov.value2 = "Realism/Romance",
  main = "Comparing Vocabulary: Shelley vs. Romance",
  plabels = c("Topic 33", "Topic 18"), 
  n = 25 
)



# "Shelley" vs. "Gothic Horror"
# pdf("Shelley vs Gothic.pdf", width = 12, height = 6) 

plot(
  stm_model, 
  type = "perspectives", 
  topics = c(21,15),  
  model = stm_model, 
  cov.value1 = "Shelley",
  cov.value2 = "Gothic Horror",
  main = "Comparing Vocabulary: Shelley vs. Gothic",
  plabels = c("Topic 19", "Topic 10"), 
  n = 25 
)

#dev.off()

# "Shelley" vs. "Sci-Fi"
#pdf("Shelley vs SciFi.pdf", width = 12, height = 6) 

plot(
  stm_model, 
  type = "perspectives", 
  topics = c(3,14),  
  model = stm_model, 
  cov.value1 = "Shelley",
  cov.value2 = "Sci-Fi",
  main = "Comparing Vocabulary: Shelley vs. Science Fiction",
  plabels = c("Topic 30", "Topic 29"), 
  n = 25 
)

#dev.off()
```

This next visualisation was more for a nice simple summary, it is a word cloud separated into each genre (by colour), highlighting the difference in vocabulary between the 3 genres and Shelley. This does look interesting, an serves as a useful summary image, or to get an idea of the difference in genres to begin analysis.

```{r comparisonwordcloud}
# Aggregate Text by Genre
genre_text = balanced_df %>%
  group_by(genre) %>%
  summarize(full_text = paste(text_segment, collapse = " "))

# Create a Term-Document Matrix
genre_corpus = VCorpus(VectorSource(genre_text$full_text))

# Clean data
genre_tdm = TermDocumentMatrix(genre_corpus, control = list(
  removePunctuation = TRUE,
  stopwords = c(stopwords("english"), my_stopwords),
  removeNumbers = TRUE,
  tolower = TRUE
))

# Convert to Matrix and Name the Columns
genre_mat = as.matrix(genre_tdm)
colnames(genre_mat) = genre_text$genre

# --- Generate the Comparison Cloud ---
cloud_colors = c(
  genre_colors["Gothic Horror"],
  genre_colors["Realism/Romance"],
  genre_colors["Sci-Fi"],
  genre_colors["Shelley"]
)


comparison.cloud(
  genre_mat,
  max.words = 200,        # Total words to show
  random.order = FALSE,   # Group them by genre
  colors = cloud_colors,  # Use your Red/Purple/Blue/Yellow palette
  title.size = 1.5,       # Size of the genre labels
  scale = c(3, 0.5)       # Size of biggest vs smallest words
)
```



# Unused Code

If analysis of era was required, the results from this were very much as expected, and therefore was ommitted from analysis.

```{r unused1}
# # --- Era Visualisation ---
# era_effect = estimateEffect(
#   1:optimal_k ~ era, 
#   stm_model, 
#   meta = out$meta)
# 
# # "Romantic" vs. "Victorian"
# plot(
#   era_effect, 
#   covariate = "era", 
#   topics = c(1:optimal_k), 
#   model = stm_model, 
#   method = "difference",
#   cov.value1 = "Late Victorian (1880-1920)",  
#   cov.value2 = "Romantic/Mid-Victorian (Pre-1880)", 
#   xlab = "More Romantic <--> More Victorian",
#   main = "Romantic vs Victorian"
# )
# 
# # "Victorian" vs "Modern"
# plot(
#   era_effect, 
#   covariate = "era", 
#   topics = c(1:optimal_k), 
#   model = stm_model, 
#   method = "difference",
#   cov.value1 = "Modern (Atomic Age)",  
#   cov.value2 = "Late Victorian (1880-1920)", 
#   xlab = "More Victorian <--> More Modern",
#   main = "Victorian vs Modern"
# )
# 
# # "Romantic" vs. "Modern"
# plot(
#   era_effect, 
#   covariate = "era", 
#   topics = c(1:optimal_k),
#   model = stm_model, 
#   method = "difference",
#   cov.value1 = "Modern (Atomic Age)",
#   cov.value2 = "Romantic/Mid-Victorian (Pre-1880)",
#   xlab = "More Romantic <--> More Modern",
#   main = "Romantic vs Modern"
# )
```

This was to consider a topic and its prevalence across different genres, however, the previous visualisations was more effective.

```{r unsused2}
# # New model comparing genre
# stm_genre_model = stm(
#   documents = out$documents, 
#   vocab = out$vocab,
#   K = 20,
#   prevalence = ~ genre + era,
#   content = ~ genre,
#   data = out$meta,
#   max.em.its = 150,
#   init.type = "Spectral"
# )
# 
# # --- Genre Visualisation ---
# # "Shelley" vs. "Realism/Romance"
# plot(
#   stm_model, 
#   type = "perspectives", 
#   topics = 7,  
#   covar = "genre", 
#   model = stm_model, 
#   cov.value1 = "Shelley",
#   cov.value2 = "Realism/Romance",
#   main = "Word Usage: Shelley (Left) vs. Sci-Fi (Right)",
#   plabels = c("Shelley", "Realism/Romance"), 
#   n = 25 
# )
# 
# # "Shelley" vs. "Gothic Horror"
# plot(
#   stm_model, 
#   type = "perspectives", 
#   topics = 7,  
#   covar = "genre", 
#   model = stm_model, 
#   cov.value1 = "Shelley",
#   cov.value2 = "Gothic Horror",
#   main = "Word Usage: Shelley (Left) vs. Sci-Fi (Right)",
#   plabels = c("Shelley", "Gothic Horror"), 
#   n = 25 
# )
# 
# # "Shelley" vs. "Sci-Fi"
# plot(
#   stm_model, 
#   type = "perspectives", 
#   topics = 7,  
#   covar = "genre", 
#   model = stm_model, 
#   cov.value1 = "Shelley",
#   cov.value2 = "Sci-Fi",
#   main = "Word Usage: Shelley (Left) vs. Sci-Fi (Right)",
#   plabels = c("Shelley", "Sci-Fi"), 
#   n = 25 
# )
```



# References
[Dealban, 2025] Dealban, D. (2025). Learning stm. GitHub Repository.

[Haldane, 2015] Haldane, M. (2015). What Frankenstein’s creature can really tell us about AI.

[Mitra, 2011] Mitra, Z. (2011). A science fiction in a gothic scaffold: A reading of mary shelley’s Frankenstein. Rupkatha
Journal on Interdisciplinary Studies in Humanities, 3(1):52–59.

[Paley, 1993] Paley, M. D. (1993). Mary shelley’s The Last Man: Apocalypse without millennium. Keats-Shelley Review,
4(1):1–25.

[Project Gutenberg, 2025] Project Gutenberg (2025). Project gutenberg. Digital Library of Free eBooks.

[Roberts et al., 2019] Roberts, M. E., Stewart, B. M., and Tingley, D. (2019). stm: An R package for structural topic
models. Journal of Statistical Software, 91(2):1–40.

[Robinson, 2025] Robinson, D. (2025). gutenbergr: Download and Process Public Domain Works from Project Gutenberg.
R package version 0.2.4.

[Silge, 2018a] Silge, J. (2018a). Evaluating STM.

[Silge, 2018b] Silge, J. (2018b). Sherlock holmes & STM.

[Silge and Robinson, 2017] Silge, J. and Robinson, D. (2017). Text Mining with R: A Tidy Approach. O’Reilly Media,
Sebastopol, CA.

[Stableford, 1995] Stableford, B. (1995). Frankenstein and the origins of science fiction. In Seed, D., editor, Anticipations:
Essays on Early Science Fiction and its Precursors. Syracuse University Press, Syracuse, NY.

[Wikipedia Contributors, 2025a] Wikipedia Contributors (2025a). Armageddon 2419 a.d. Accessed: 2025-12-07.

[Wikipedia Contributors, 2025b] Wikipedia Contributors (2025b). Gothic fiction. Accessed: 2025-12-07.

# AI Usage Report
AI Usage Reporting: Transparency Statement

This project utilised Large Language Models (LLMs) to assist in the technical implementation and editorial refinement of
the research. The specific contributions of the AI are detailed below:

1. Code Development and Debugging
The R code used for data acquisition, preprocessing, and modelling was co-developed with AI assistance. Specifically, the

AI was used to:
- Troubleshoot errors in the gutenbergr package mirror configuration.
- Generate the syntax for the stm package, specifically regarding the estimateEffect and topicCorr functions.
- Draft the ggplot2 code for the custom visualisation of topic prevalence and vocabulary comparison clouds.
- Verification: All code was executed, tested, and validated by the author on a local machine to ensure reproducibility
and accuracy.

2. Data Structuring

The AI assisted in compiling the list of Project Gutenberg IDs for the comparative corpus (Romantic, Gothic, and Science

Fiction novels) to ensure a balanced dataset.

It generated the LaTeX table code to format the raw results into academic-standard tables (e.g., Table 11: Thematic
Super-Clusters).

3. Editorial and Stylistic Refinement

Draft sections of the report were submitted to the AI for proofreading. The AI provided suggestions for correcting spelling,
grammar, and academic tone.

The AI assisted in refining the "Problem Statement" and "Abstract" to ensure conciseness and clarity.

 Note: The core literary arguments, the interpretation of the "Bridge" hypothesis, and the final conclusions regarding Shelley’s
technological limitations are the original intellectual work of the author. The AI functioned solely as an editor and technical
assistant.

4. Bibliographic Formatting

The AI was used to convert raw URLs and citations into formatted BibTeX entries to ensure compatibility with the LaTeX
document structure.



